Exploring mathematical objects from custom-tailored mathematical universes

### 2020-01-18 paper-filmat.tex
to do

\begin{tabular}{lll}
  \toprule
  Statement & in $\Set$ & in $\Eff$ \\
  \midrule
  Any natural number is prime or not prime. & \cmark{} (trivially so) & \cmark \\
  There are infinitely many primes. & \cmark & \cmark \\
  Any function $\NN \to \NN$ is the zero function or not. & \cmark{} (trivially so) & \xmark \\
  Any function $\NN \to \NN$ is computable by a Turing machine. & \xmark & \cmark{} (trivially so) \\
  Any function $\RR \to \RR$ is continuous. & \xmark & \cmark \\
  \bottomrule
\end{tabular}

@@

* Cauchy vs. Dedekind numbers (physical quantities, ...)
* Syntactical vs. semantical interpretation
* phone call analogy
* detailed explanation of the pretty picture; analogy with "inner models"
  of set theory
* quick overview of the several aspects of toposes (maybe at the end,
  as an outlook?)
* allowing a switch of perspective
* applications in mathematical practice
* int. logic as common denominator
* arbitrariness of the "standard axioms"
* what giving up classical logic actually amounts to in practice
* importance of having an adapted language
* SDG (infinitesimals are well-studied in philosophy of mathematics)
* quick remark on the internal language being based on types instead of sets
* Source https://plato.stanford.edu/entries/category-theory/ for relevant
  literatur
* the internal language as a generalized modal operator

* use the phrase "very rich and very deep" (or similarly) about set theory
* PA doesn't have models
* Sven(?)
* semantic view vs. syntactic view
* Kripke models
* more pointers to topos literature
* Reference for the effective topos
* Reference idempotency
* A topos is more than its semantics
* Reference: Moerdijk
* Check references everywhere (give good pointers)
* Format rules in a better way
* Format examples in Eff in a better way
* toposes are not small (Figure 1)

@@

Mention forcing

>>>>

\begin{tabular}{lll}
  \toprule
  Statement & in $\Set$ & in $\Eff$ \\
  \midrule
  Any natural number is prime or not prime. & \cmark{} (trivially so) & \cmark \\
  There are infinitely many primes. & \cmark & \cmark \\
  Any function $\NN \to \NN$ is the zero function or not. & \cmark{} (trivially so) & \xmark \\
  Any function $\NN \to \NN$ is computable by a Turing machine. & \xmark & \cmark{} (trivially so) \\
  Any function $\RR \to \RR$ is continuous. & \xmark & \cmark \\
  Heyting arithmetic is categorical. & \xmark & \cmark \\
  \bottomrule
\end{tabular}

@@

* XXX Cauchy vs. Dedekind numbers (physical quantities, ...)
* XXX Syntactical vs. semantical interpretation
* XXX phone call analogy
* XXX detailed explanation of the pretty picture; analogy with "inner models"
  of set theory
* XXX quick overview of the several aspects of toposes (maybe at the end,
  as an outlook?)
* XXX allowing a switch of perspective
* XXX applications in mathematical practice
* XXX int. logic as common denominator
* XXX arbitrariness of the "standard axioms"
* XXX what giving up classical logic actually amounts to in practice
* XXX importance of having an adapted language
* XXX SDG (infinitesimals are well-studied in philosophy of mathematics)
* XXX quick remark on the internal language being based on types instead of sets
* XXX Source https://plato.stanford.edu/entries/category-theory/ for relevant
  literatur
* XXX the internal language as a generalized modal operator

* XXX use the phrase "very rich and very deep" (or similarly) about set theory
* XXX PA doesn't have models
* XXX Sven(?)
* XXX semantic view vs. syntactic view
* XXX Kripke models
* XXX more pointers to topos literature
* XXX Reference for the effective topos
* XXX Reference idempotency
* XXX A topos is more than its semantics
* XXX Reference: Moerdijk
* XXX Check references everywhere (give good pointers)
* XXX Format rules in a better way
* XXX Format examples in Eff in a better way
* XXX toposes are not small (Figure 1)

@@
